### 2024-11-14 paper/paper.qmd
Add files via upload

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

---
title: "Analysis of Hate Crime Trends in the United States"
author: "Yuxuan Wei"
date: "12 November 2024"
format: pdf
---

# Abstract

This study analyzes trends in hate crimes across the United States, examining the influence of various socioeconomic factors. Using data from FiveThirtyEight, we model hate crime rates over time to provide insights into the distribution and intensity of hate crimes across regions. Our findings contribute to the understanding of the social dynamics that impact hate crime occurrences, offering actionable insights for policymakers and communities.

# Introduction

Hate crimes are a significant social issue with profound implications for public safety and social cohesion. This paper investigates hate crime trends across the United States and explores the relationship between hate crime rates and socioeconomic variables such as median income, unemployment rates, and population density. By examining these factors, this paper seeks to identify key drivers of hate crimes and assess the predictive power of these variables on hate crime rates.

The structure of this paper is as follows: Section 2 discusses the data and measurement considerations, Section 3 details the modeling approach, Section 4 presents the results, and Section 5 provides a discussion and conclusion.

# Data

## Overview

The primary data source for this analysis is the hate crimes dataset provided by FiveThirtyEight, which contains information on hate crime occurrences, economic indicators, and demographic details for each U.S. state. Key variables include:
- **Hate Crime Rate**: The number of hate crimes per 100,000 people.
- **Median Income**: Median household income by state.
- **Unemployment Rate**: State-level unemployment rate.
- **Share of Population in Metro Areas**: The proportion of the population living in metropolitan areas.

Data were cleaned and formatted to facilitate analysis, and missing values were addressed using appropriate imputation techniques where necessary.

## Measurement and Limitations

The hate crime rate is measured as the number of reported hate crimes per 100,000 people, which provides a normalized view of hate crime prevalence across states. However, limitations arise due to reporting inconsistencies and potential underreporting, which may impact the accuracy of the data.

# Methodology

## Model Selection

To investigate the impact of socioeconomic factors on hate crime rates, a multiple linear regression model was chosen. This model allows us to quantify the effect of each predictor variable while controlling for other factors. The regression model is defined as follows:

\[
\text{Hate Crime Rate} = \beta_0 + \beta_1 \text{Median Income} + \beta_2 \text{Unemployment Rate} + \beta_3 \text{Share of Population in Metro} + \epsilon
\]

where:
- \(\beta_0\) is the intercept,
- \(\beta_1\), \(\beta_2\), and \(\beta_3\) are the coefficients for each predictor variable,
- \(\epsilon\) is the error term.

The model assumptions, including linearity, independence, homoscedasticity, and normality of residuals, were checked and verified using diagnostic plots.

## Alternative Models Considered

Although multiple linear regression was selected, alternative models such as logistic regression and Poisson regression were considered but found unsuitable due to the continuous nature of the outcome variable.

# Results

## Summary Statistics

Table 1 provides descriptive statistics for key variables in the dataset.

```{r}
# Load necessary libraries and data
library(tidyverse)

# Display summary statistics
data <- read.csv("cleaned_hate_crimes.csv")
summary(data)